---
title: "B209415"
output: pdf_document
GIThub: https://github.com/B209415/B209415_assessment
---
The data used is test data from England accident and emergency attendances and admissions (ae_attendances_test) in CSV format.
#It contains the following variables
Reported attendances, four-hour breaches and admissions for all A&E departments in England for the years 2016/17 through 2018/19 (Apr-Mar). which is part  NHS data in CSV format.
#Tool generation and data collection
The tool that used for data collection from ae attendances test data was created using python injupyter. interactive widgets was loaded and used for tool creation.The codes on python was modified from 'Collecting data using interactive Jupyter widgets.ipynb'  Jupyter Notebook

#New R markdown file was created 
based on 'Loading NHSRdatasets.Rmd'. It contains: link to the GIThub repository,  Annotations that expain each step and code  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## loadin pachages

Firstly we will need to load all needed packages 

```{r cars}
library(tidyverse)
library(readr)
library(here)
library(knitr)
library(scales)
library(lubridate)
library(caret)

```
#Loading NHS data
Here we will load NHS data used for our coding, data collection and analysis

```{r cars}
library(NHSRdatasets)
library(readr)
```

```
#Loading the data collected by python tool
CD is abbreviation for created data, it will be used as a shortcut to call for the data


```{r}

CD <- read_csv("~/B209415/Data/createdData.csv")
View(CD)

```
```
#Know the class of the all data together
The data is classified as tibble
```{r}
class(CD)
```
```
#Look at the data
This will show the data as well as the type of the data which important for analysis
Data types in the collected data are: Index, attendances,breaches (decimal:numeric), period (date), consent (logical)
All the data type are in the needed category so no need to reclass them


```{r}
glimpse(CD)

```

```
#Data dictionary
Here we will install and load the packages needed for creation of data dictionary

```{r}
install.packages('dataMeta')
library(dataMeta)

```

#Now lets look at the data through glimpse funtion
It give information on the length/type of data
The data had 5 coloumns (variables), 11 raws (observations)
The type of the variable are: index, attendences, breaches (decimal), period (date/numeric), consent (logical)
```{r}
glimpse(CD)
```
#here we will build and descripe our data frame

```{r}
variable_description <- c("The index column that allows us to link the data collected to the original ae_attendances data in the 'RawData' folder.",
"The month that this activity relates to, stored as a date (1st of each month).",
"The number of attendances for this department type at this organisation for this month.", 
"The number of attendances that breached the four-hour target.",
"The consent from the end-user to process and share the data collected with the data capture tool.")

print(variable_description)

```


```{r}
variable_type <- c(0, 1, 0, 0, 1)
print(variable_type)
```
# Now using the build_linker funtion the data frame, variable discription and variable type will be linked into data dictionary

```{r}
linker<-build_linker(CD, variable_description, variable_type)
print(linker)
dictionary <- build_dict(my.data = CD, linker = linker)
glimpse(dictionary)

```
now we will save our dictionary

```{r}
write_csv(dictionary, here("RawData", "CD_DD.csv"))

```

#Now the exploratory data analysis will start
Summary statistics for attendances and breaches
Mean, Standard deviation, Minimum and Maximum value


```{r}
#Table:summerize the attendences data
tableA <- CD %>%
drop_na(attendances) %>%
  summarize(meanattendances = mean(attendances)
            , SDattendances = sd(attendances), 
            minattendances = min(attendances), 
            maxattendances = max(attendances))
tableA

#tableB: summerize the breaches data
tableB <- CD %>%
  drop_na(attendances) %>%
  summarize(meanbreaches = mean(breaches),
            SDbreaches = sd(breaches), 
            minbreaches = min(breaches), 
            maxbreaches = max(breaches))
tableB

```

#Showing the summary of attendences and breaches by year

```{r}
#TableC: shows the summary statstics for attendences by year
tableC <- CD %>%
  mutate_at(vars(period), format, "%y") %>%
  group_by(period) %>%
  summarize(meanattendances = mean(attendances),
            SDattendances = sd(attendances), 
            sumattendences = sum(attendances),
            minattendances = min(attendances), 
            maxattendances = max(attendances))
tableC

#TableC: shows the summary statstics for breaches by year
tableD <- CD %>%
  mutate_at(vars(period), format, "%y") %>%
  group_by(period) %>%
  summarize(meanbreaches = mean(breaches),
            SDbreaches = sd(breaches), 
            sumbreaches = sum(breaches),
            minbreaches = min(breaches), 
            maxbreaches = max(breaches))
tableD
```

```
#Creation of graphs that shows the the attendances and breaches by year
The period is in the x-axis, attendences in the y-axis, whie the size of the puple represent the number of breaches amonge attendences each year
```{r pressure, echo=FALSE}
plot(pressure)
graph_B <- CD %>%
  ggplot() +
  geom_point(aes(period, attendances, size = breaches))
graph_B
```


